% Ch5 — BV Functions and Sets of Finite Perimeter
% Evans & Gariepy pp. 166–226. This is the hardest chapter in the book.

\section{October 21, 2024}\label{sec:oct21}

Midterm grades came back. Could have been worse. Today we start Chapter 5: BV functions and sets of finite perimeter. Evans pp.\ 166--226 is one of the hardest stretches in the book, and I can feel it already from the density of the notation alone.

The whole point of BV is to handle jumps that $W^{1,1}$ cannot. Take the characteristic function of a ball. It has a perfectly good ``derivative'' in the distributional sense: a surface measure on the boundary. But $\chi_E \notin W^{1,1}$ because that derivative is not an $L^1$ function. BV says: fine, let the derivative be a measure. That is the entire motivation. I'm drinking lukewarm coffee from the thermos I forgot to wash out yesterday, and it tastes about as good as you'd expect.

\subsection{Definitions and first properties}\label{subsec:bv-defs}

Let $U \subseteq \R^n$ be open.

\begin{definition}[Variation]\label{def:bv-variation}
For $f \in L^1(U)$, define the \emph{variation} of $f$ in $U$ by
\[
  \norm{Df}(U) \;=\; \sup\!\left\{ \int_U f \,\operatorname{div}\varphi \, dx \;\bigg|\; \varphi \in C_c^1(U; \R^n),\; |\varphi| \le 1 \right\}.
\]
Say $f \in BV(U)$ if $\norm{Df}(U) < \infty$.
\end{definition}

Test $f$ against all smooth compactly supported divergence fields of magnitude at most 1 and take the sup. If finite, you are in $BV$. The sup is the total variation of $Df$, which by the Riesz representation theorem (the one from Chapter~1, Theorem~\ref{thm:riesz}) is an $\R^n$-valued Radon measure.

The norm on $BV(U)$ is $\norm{f}_{BV} = \norm{f}_{L^1} + \norm{Df}(U)$.

\begin{remark}\label{rmk:sobolev-subset-bv}
$W^{1,1}(U) \subset BV(U)$. If $f \in W^{1,1}$, then $Df = \nabla f \, dx$ with $\nabla f \in L^1$, and $\norm{Df}(U) = \int_U |\nabla f| \, dx < \infty$. But BV is strictly bigger: $\chi_E$ for $E$ with smooth boundary is in $BV$ but not $W^{1,1}$, because its derivative is a surface measure on $\partial E$.
\end{remark}

\begin{example}\label{ex:bv-canonical}
Some examples worth keeping in your head:
\begin{enumerate}
  \item $f = \chi_E$ where $E$ is a ball in $\R^n$. Then $f \in BV(\R^n)$ and $\norm{Df}(\R^n) = \mathcal{H}^{n-1}(\partial E)$. The derivative is $(-\nu_E) \mathcal{H}^{n-1} \lfloor \partial E$ where $\nu_E$ is the outward unit normal.
  \item The Cantor function on $[0,1]$: in $BV$ but not $W^{1,1}$. Its derivative is entirely singular, concentrated on the Cantor set. No jump part either. All Cantor part.
  \item Any monotone function $f: \R \to \R$ is $BV_{\text{loc}}$. Connects to the classical theory of functions of bounded variation on the real line.
\end{enumerate}
\end{example}

\subsection{The structure theorem}\label{subsec:structure}

This makes everything else in the chapter go.

\begin{theorem}[Structure Theorem for BV]\label{thm:structure-bv}
Let $f \in BV(U)$. Then $Df$ is an $\R^n$-valued Radon measure on $U$, admitting the decomposition
\[
  Df \;=\; D^a f + D^s f,
\]
where:
\begin{itemize}
  \item $D^a f$ is absolutely continuous with respect to $\mathcal{L}^n$. By Radon--Nikodym, $D^a f = \nabla f \, \mathcal{L}^n$ for some $\nabla f \in L^1(U;\R^n)$.
  \item $D^s f$ is singular with respect to $\mathcal{L}^n$: concentrated on a set of Lebesgue measure zero.
\end{itemize}
\end{theorem}

This is Lebesgue--Radon--Nikodym decomposition applied to the measure $Df$. Nothing deep yet, but naming the pieces is what lets you talk precisely about the structure.

For $\chi_E$, the absolutely continuous part vanishes and the singular part is the surface measure on $\partial E$. For a smooth function, it is all absolutely continuous. For the devil's staircase, it is all singular.

$D^s f$ decomposes further into $D^j f + D^c f$: a \emph{jump part} concentrated on a countably $(n-1)$-rectifiable set and a \emph{Cantor part} that is singular but diffuse. We will need this in the pointwise properties section later.

\subsection{Approximation and compactness}\label{subsec:approx-compact}

Two big results. First, lower semicontinuity.

\begin{theorem}[Lower semicontinuity]\label{thm:lsc-variation}
If $f_j \to f$ in $L^1(U)$, then
\[
  \norm{Df}(U) \;\le\; \liminf_{j \to \infty} \norm{Df_j}(U).
\]
\end{theorem}

Why? $\norm{Df}(U)$ is a supremum of continuous linear functionals of $f$. Each functional passes to the limit, and the sup of limits is $\le$ the liminf of sups. I had to sit down and write it out to believe it. Standard argument, but I keep getting the quantifiers backward until I write it slowly.

\begin{theorem}[Approximation by smooth functions]\label{thm:approx-bv}
If $f \in BV(U)$, there exist $f_j \in C^\infty(U) \cap BV(U)$ with
\[
  f_j \to f \;\text{in } L^1(U) \quad\text{and}\quad \norm{Df_j}(U) \to \norm{Df}(U).
\]
\end{theorem}

Here is the catch. This is convergence in the \emph{strict} topology: $L^1$ convergence of functions plus convergence of total variation. You do NOT get $f_j \to f$ in the BV norm. That would require $\norm{Df_j - Df}(U) \to 0$, which is impossible in general. Why? BV functions can have singular parts and smooth functions cannot. The mollification $f * \rho_\varepsilon$ converges in $L^1$ and the variation converges, but $D(f * \rho_\varepsilon) - Df$ does not go to zero as a measure. I confused this with norm convergence for two weeks.

BV is not the closure of $C^\infty$ in the BV norm. That closure is $W^{1,1}$.

\begin{theorem}[Compactness]\label{thm:compactness-bv}
Let $\{f_j\}$ be a sequence in $BV(U)$ with $\sup_j \norm{f_j}_{BV(U)} < \infty$, where $U$ is bounded with Lipschitz boundary. Then a subsequence converges in $L^1(U)$ to some $f \in BV(U)$.
\end{theorem}

In Sobolev spaces you would get weak convergence by reflexivity, but $W^{1,1}$ is not reflexive and neither is $BV$. The compactness is only in $L^1$. Combined with lower semicontinuity, it is enough for the direct method in the calculus of variations. That combination (compactness plus l.s.c.) shows up so often that I should probably tattoo it on my arm.

\subsection{Traces and extensions}\label{subsec:trace-ext}

\begin{theorem}[Trace for BV]\label{thm:trace-bv}
Let $U \subset \R^n$ be bounded open with Lipschitz boundary. There's a bounded linear operator
\[
  T: BV(U) \to L^1(\partial U; \mathcal{H}^{n-1})
\]
with $Tf = f|_{\partial U}$ for $f \in BV(U) \cap C(\overline{U})$.
\end{theorem}

The trace only lands in $L^1(\partial U)$. No fractional Sobolev regularity. Makes sense: BV functions can jump, so the trace just records the boundary values. The proof uses the same flattening machinery as Sobolev traces but needs the BV-specific approximation results from above.

\begin{theorem}[Extension]\label{thm:extension-bv}
Let $U \subset \R^n$ be bounded with Lipschitz boundary. There's a bounded linear operator $E: BV(U) \to BV(\R^n)$ with $Ef = f$ a.e.\ in $U$, $\spt(Ef)$ compact, and $\norm{Ef}_{BV(\R^n)} \le C \norm{f}_{BV(U)}$.
\end{theorem}

Mirrors the Sobolev case. The Lipschitz boundary condition is doing real work.


\section{October 23, 2024}\label{sec:oct23}

Today: coarea formula for BV and isoperimetric inequality. I think the coarea formula is the single best result in the chapter. Brian said the same thing last week, independently, which either means we are both right or we have been talking to each other too much.

\subsection{Coarea formula for BV}\label{subsec:coarea}

\begin{theorem}[Coarea formula]\label{thm:coarea-bv}
Let $f \in BV(U)$. For $\mathcal{L}^1$-a.e.\ $t \in \R$, the level set $\{f > t\}$ has finite perimeter in $U$. And for every Borel set $A \subseteq U$,
\[
  \norm{Df}(A) \;=\; \int_{-\infty}^{\infty} \norm{\partial\{f > t\}}(A)\, dt,
\]
where $\norm{\partial\{f > t\}}$ is the perimeter measure of the superlevel set $\{x \in U : f(x) > t\}$.
\end{theorem}

Let me unpack this. The left side is the total variation of $Df$ on $A$. The right side slices this into contributions from each level: at height $t$, the superlevel set $\{f > t\}$ contributes its perimeter restricted to $A$. Integrate over all heights and you recover the total variation.

For $f \in W^{1,1}(U)$, the formula becomes
\[
  \int_A |\nabla f| \, dx \;=\; \int_{-\infty}^{\infty} \mathcal{H}^{n-1}(A \cap \partial^*\{f > t\})\, dt,
\]
which generalizes the classical coarea formula from Chapter~3.

\begin{remark}\label{rmk:coarea-converse}
The formula also gives a converse: $f \in L^1(U)$ is in $BV(U)$ if and only if $\int_{-\infty}^\infty \norm{\partial\{f > t\}}(U)\,dt < \infty$. Sometimes that is the easiest way to check BV membership. I used it on a problem set for a radial function and it was far cleaner than the definition.
\end{remark}

The proof confused me on first read because Evans reduces to smooth $f$ first, proves the smooth case by the classical coarea formula, then passes to the limit using lower semicontinuity. The $\le$ and $\ge$ directions are proved separately. The $\ge$ direction is harder.

\subsection{Isoperimetric inequalities}\label{subsec:isoperimetric}

Among all sets of given volume, the ball has the smallest perimeter. Known since antiquity, which makes you wonder what the ancient Greeks would have done with distributional derivatives.

\begin{theorem}[Isoperimetric inequality]\label{thm:isoperimetric}
Let $E \subset \R^n$ have finite perimeter. Then
\[
  \norm{\partial E}(\R^n) \;\ge\; n\,\alpha(n)^{1/n}\, \bigl(\mathcal{L}^n(E)\bigr)^{(n-1)/n},
\]
where $\alpha(n) = \mathcal{L}^n(B(0,1))$. Equality iff $E$ is a ball (up to measure zero).
\end{theorem}

The argument uses the coarea formula. First establish the BV Sobolev inequality: for $f \in BV(\R^n)$,
\[
  \norm{f}_{L^{n/(n-1)}(\R^n)} \;\le\; \frac{1}{n\,\alpha(n)^{1/n}} \norm{Df}(\R^n).
\]
Then plug in $f = \chi_E$. The left side gives $\mathcal{L}^n(E)^{(n-1)/n}$ and the right gives $\norm{\partial E}(\R^n)$ times the constant.

Actually, that is only the inequality. The equality case requires a symmetrization argument, which is a different beast entirely. But the inequality itself follows from the coarea formula in about half a page.

There is also a \emph{relative isoperimetric inequality}: if $U$ is bounded Lipschitz and $\mathcal{L}^n(E \cap U) \le \frac{1}{2}\mathcal{L}^n(U)$, then
\[
  \mathcal{L}^n(E \cap U)^{(n-1)/n} \;\le\; C(U)\, \norm{\partial E}(U).
\]
Useful for compactness arguments.


\section{October 28, 2024}\label{sec:oct28}

This chapter is harder than Chapter 4. I am sure of it now. Today we get into the geometric measure theory proper: reduced boundary, blow-up, and the generalized divergence theorem. I sat in the front row for this one.

\subsection{Sets of finite perimeter and the reduced boundary}\label{subsec:reduced-boundary}

For a set $E$ of finite perimeter, the ``boundary'' in the BV sense is not the topological boundary $\partial E$ (which can be wild). It is a well-behaved subset called the \emph{reduced boundary} $\partial^* E$.

\begin{definition}[Reduced boundary]\label{def:reduced-boundary}
Let $E$ have finite perimeter in $\R^n$. Write $\mu_E = D\chi_E$ for the Gauss--Green measure and $\norm{\mu_E} = \norm{\partial E}$ for its total variation. A point $x \in \R^n$ belongs to the \emph{reduced boundary} $\partial^* E$ if:
\begin{enumerate}
  \item $\norm{\mu_E}(B(x,r)) > 0$ for all $r > 0$.
  \item The limit $\nu_E(x) = \lim_{r \to 0} \frac{\mu_E(B(x,r))}{\norm{\mu_E}(B(x,r))}$ exists.
  \item $|\nu_E(x)| = 1$.
\end{enumerate}
The vector $\nu_E(x)$ is called the \emph{measure-theoretic outer unit normal}.
\end{definition}

$\partial^* E$ consists of points where the boundary ``looks like a hyperplane'' at small scales. At each such point there is a meaningful notion of which way is out. The reduced boundary sits inside the topological boundary $\partial E$, and the two can differ wildly. Modify $E$ on a null set and the topological boundary can change beyond recognition, but $\partial^* E$ does not move.

\subsection{The blow-up theorem}\label{subsec:blowup}

This is the hard theorem of the chapter. I do not have a clean one-paragraph summary for it.

\begin{theorem}[Blow-up at the reduced boundary]\label{thm:blowup}
Let $E$ have finite perimeter, $x \in \partial^* E$ with outer normal $\nu_E(x)$. Define
\[
  E_{x,r} \;=\; \frac{E - x}{r}.
\]
Then $\chi_{E_{x,r}} \to \chi_{H^-(\nu_E(x))}$ in $L^1_{\mathrm{loc}}(\R^n)$ as $r \to 0^+$, where $H^-(\nu) = \{y : y \cdot \nu < 0\}$.
\end{theorem}

Zoom in at a point of $\partial^* E$ and the set converges to a half-space. The boundary looks flat at infinitesimal scales. The normal to the half-space is $\nu_E(x)$ from the definition. The proof uses Besicovitch differentiation, the definition of the reduced boundary, and BV compactness. Evans pp.\ 190--200. I went through it once and got the main ideas. Reproducing it from scratch would take me a full afternoon.

\begin{remark}\label{rmk:rectifiability}
A major consequence: $\partial^* E$ is countably $(n-1)$-rectifiable, and the perimeter measure satisfies $\norm{\partial E} = \mathcal{H}^{n-1} \lfloor \partial^* E$. The perimeter measure is $(n-1)$-dimensional Hausdorff measure restricted to the reduced boundary.
\end{remark}

\subsection{Measure-theoretic boundary and Gauss--Green}\label{subsec:gauss-green}

\begin{definition}[Measure-theoretic boundary]\label{def:mt-boundary}
The \emph{measure-theoretic boundary} $\partial_* E$ consists of all $x \in \R^n$ where neither $E$ nor its complement has density 1:
\[
  \partial_* E \;=\; \left\{x : \limsup_{r \to 0} \frac{\mathcal{L}^n(E \cap B(x,r))}{\alpha(n)r^n} > 0 \;\text{and}\; \limsup_{r \to 0} \frac{\mathcal{L}^n(B(x,r) \setminus E)}{\alpha(n)r^n} > 0\right\}.
\]
\end{definition}

We always have $\partial^* E \subseteq \partial_* E \subseteq \partial E$. For sets of finite perimeter, $\mathcal{H}^{n-1}(\partial_* E \setminus \partial^* E) = 0$, so the reduced and measure-theoretic boundaries agree up to $\mathcal{H}^{n-1}$-null sets.

Now the payoff.

\begin{theorem}[Generalized Gauss--Green theorem]\label{thm:gauss-green}
Let $E \subset \R^n$ have finite perimeter, $\varphi \in C_c^1(\R^n; \R^n)$. Then
\[
  \int_E \operatorname{div}\varphi\, dx \;=\; \int_{\partial^* E} \varphi \cdot \nu_E \, d\mathcal{H}^{n-1}.
\]
\end{theorem}

The classical divergence theorem needs $E$ to have smooth or at least Lipschitz boundary. This version only needs finite perimeter. The boundary integral is over $\partial^* E$ with the measure-theoretic normal. The cleanest way to see why: $D\chi_E = -\nu_E \mathcal{H}^{n-1} \lfloor \partial^* E$ by the structure theory, and the formula is just the definition of distributional derivative applied to $\chi_E$. Once you have the blow-up theorem and rectifiability, this falls out.

\subsection{Pointwise properties of BV functions}\label{subsec:pointwise}

What can you say about individual points? More than I expected.

\begin{theorem}[Approximate limits and jump set]\label{thm:pointwise-bv}
Let $f \in BV(U)$. Then:
\begin{enumerate}
  \item For $\mathcal{H}^{n-1}$-a.e.\ $x \in U$, $f$ has an approximate limit $\tilde{f}(x) = \operatorname{ap\,lim}_{y \to x} f(y)$ (the precise representative).
  \item The jump set $J_f$ is countably $(n-1)$-rectifiable.
  \item $D^s f = D^j f + D^c f$, where $D^j f = (f^+ - f^-) \nu_f \, \mathcal{H}^{n-1} \lfloor J_f$ is the jump part and $D^c f$ is the Cantor part.
\end{enumerate}
\end{theorem}

So the full decomposition is:
\[
  Df \;=\; \underbrace{\nabla f \, \mathcal{L}^n}_{\text{abs.\ cont.}} \;+\; \underbrace{(f^+ - f^-)\nu_f \, \mathcal{H}^{n-1}\lfloor J_f}_{\text{jump}} \;+\; \underbrace{D^c f}_{\text{Cantor}}.
\]

Three parts, each living on a different kind of set. The ac part is spread over $U$. The jump part sits on the $(n-1)$-dimensional jump set. And the Cantor part lives on something stranger, like the Cantor set itself. Three flavors of derivative, each with its own geometry. I spent an evening staring at this decomposition and thinking about which real-valued BV functions on $\R$ realize each part. Monotone + continuous + singular = all three at once, but constructing an explicit example in $\R^n$ for $n \ge 2$ seems harder.

For $\chi_E$ with $E$ having finite perimeter: $\nabla\chi_E = 0$, $D^c\chi_E = 0$, and $D^j\chi_E = -\nu_E \,\mathcal{H}^{n-1}\lfloor\partial^*E$. All jump.

\subsection{Essential variation on lines}\label{subsec:essential-variation}

\begin{theorem}[Characterization via sections]\label{thm:sections-bv}
$f \in L^1(U)$ is in $BV(U)$ if and only if for each coordinate direction $e_i$, the one-dimensional slices $t \mapsto f(x' + te_i)$ have finite essential variation for $\mathcal{L}^{n-1}$-a.e.\ $x'$, and the total of these variations is finite.
\end{theorem}

BV analogue of the ACL characterization for Sobolev functions. For Sobolev, the slices are absolutely continuous; for BV, the slices have bounded essential variation. The parallel is clean.

\subsection{Criterion for finite perimeter}\label{subsec:criterion}

\begin{theorem}[Criterion]\label{thm:criterion-perimeter}
$E \subset \R^n$ measurable has finite perimeter in $U$ iff
\[
  \sup\!\left\{ \int_E \operatorname{div}\varphi\, dx \;\bigg|\; \varphi \in C_c^1(U;\R^n),\; |\varphi| \le 1\right\} < \infty,
\]
which is $\chi_E \in BV(U)$.
\end{theorem}

Sets of finite perimeter include: sets with Lipschitz boundary, convex sets, sublevel sets $\{f < t\}$ of BV functions for a.e.\ $t$ (by the coarea formula). Sets that do NOT have finite perimeter: fractal boundaries with infinite $\mathcal{H}^{n-1}$-measure, like the Koch snowflake in $\R^2$. The Koch snowflake has finite area but its boundary has infinite length, so it fails the criterion.
